---
title: "LCC2005"
author: "Alex Chubaty & Eliot McIntire" 
date: "07 September 2016"
output: html_document
---

This is an example model of forest dynamics. It includes simple versions of forest succession in Canada's boreal forest, a simple cellular automaton fire model, some GIS data operations to select subset areas of a larger dataset, a reclassification module from a many-class land cover to a few-class land cover and a simple agent movement model. These are all toy versions of each of these types and so are not useful for real world applications; but they allow highlighting of some of SpaDES's capabilities.

This example uses the notion of a 'module group' or 'parent' module that contains several child modules:

- `caribouMovementLcc`
- `cropReprojectLccAge`
- `fireSpreadLcc`
- `forestAge`
- `forestSuccessionBeacons`
- `LccToBeaconsReclassify`

This code also shows how modules can have different time units, specifically, the child modules which have 2 different time units (month for caribou, year for others).

The code will download all modules and data required to run the spades call at the end.
If this is running on a Windows OS or on an OS X machine, the visualizations will be quick because a device will be opened that is not within RStudio.
On a Linux, this will not occur, so visualization will be slow (RStudio graphics device is slow).

```{r download-modules, eval=FALSE}
library(SpaDES)

# pick a directory to download modules and data to. Can use a temp directory
workDirectory <- file.path(dirname(tempdir()), "Lcc2005")
moduleDir <- file.path(workDirectory, "modules")
checkPath(moduleDir, create = TRUE)

downloadModule("LCC2005", moduleDir) # default `data=FALSE` doesn't download data
```


Module groups make loading multiple modules easier: only the name of the module group needs to be specified in the `simInit` call, which will then initialize the simulation with the child modules.

```{r module-group-init, eval=FALSE}
# setup simulation
outputDir <- file.path(workDirectory, "simOutputs")
cacheDir <- checkPath(file.path(outputDir, "cache"), create = TRUE)

times <- list(start = 2005.0, end = 2020.0, timeunit = "year")
parameters <- list(
  .globals = list(burnStats = "fireStats"),
  fireSpreadLcc = list(drought = 1.2), # in
  caribouMovementLcc = list(N = 1e3, startTime = times$start + 1, 
                            glmInitialTime = NA_real_)
)
modules <- list("LCC2005")
paths <- list(
  cachePath = cacheDir,
  modulePath = moduleDir,
  inputPath = moduleDir,
  outputPath = outputDir
)

# This next step will download data if they do not yet exist locally
mySim <- simInit(times = times, params = parameters, modules = modules,
                 paths = paths)

```

Now that the `mySim` object has been initialized, we can run the simulation by calling
`spades`

```{r run-SpaDES, eval=FALSE}
dev() # on Windows and Mac, this opens external device if using Rstudio, it is faster
clearPlot()

mySimOut <- spades(mySim)

# Plot Canada Map
Plot(mySimOut$lcc05, title = "Land Cover, with simulated area")
# Plot small Yellow polygon showing area simulated
Plot(mySimOut$inputMapPolygon, addTo = "mySimOut$lcc05", gp = gpar(col = "yellow", lwd = 2))

```

In addition to running a model, there are many things built into SpaDES that allow a user to explore the modules and models.

```{r Exploring model, eval = FALSE}
openModules(mySim) # this will try to open the modules in a text editor, or give instructions how to do it manually

### Simulation overview: note the child modules are initialized
moduleDiagram(mySim)
objectDiagram(mySim)

# show modules that are included in this simList
modules(mySim)

# show all objects contained within the final simList
ls(mySimOut)

# more detail
ls.str(mySimOut)

# show parameters that can easily be changed in simInit call
params(mySim)

```
