---
title: "LCC2005"
author: "Alex Chubaty"
date: "10 August 2015"
output: pdf_document
---

This is an example of a 'module group' containing several child modules:

- `caribouMovementLcc`
- `cropReprojectLccAge`
- `fireSpreadLcc`
- `forestAge`
- `forestSuccessionBeacons`
- `LccToBeaconsReclassify`

```{r model-inputs, eval=FALSE}
library(igraph)
library(raster)
library(SpaDES)

inputDir <- file.path(tempdir(), "inputs") %>% checkPath(create=TRUE)

# download data (these are large files!)
if (!file.exists(file.path(inputPath, "LandCoverOfCanada2005_V1_4", "LCC2005_V1_4a.tif"))) {
  dir.create(file.path(inputPath, "LandCoverOfCanada2005_V1_4"))
  url <- "ftp://ftp.ccrs.nrcan.gc.ca/ad/NLCCLandCover/LandcoverCanada2005_250m/LandCoverOfCanada2005_V1_4.zip"
  download.file(url = url, destfile = file.path(inputPath, "LandCoverOfCanada2005_V1_4.zip"))
  unzip(file.path(inputPath, "LandCoverOfCanada2005_V1_4.zip"))
}

if (!file.exists(file.path(inputPath, "age", "age.tif"))) {
  dir.create(file.path(inputPath, "age"))
  url <- "ftp://ftp.daac.ornl.gov/data/nacp/NA_TreeAge//data/can_age04_1km.tif"
  download.file(url = url, destfile = file.path(inputPath, "age", "age.tif"))
}

# load data
age <- raster(file.path(inputPath, "age/age.tif"))
lcc05 <- raster(file.path(inputPath, "LandCoverOfCanada2005_V1_4/LCC2005_V1_4a.tif"))
lcc05CRS <- crs(lcc05)

# Random polygon
areaKm2 <- 2000
minX <- -1072250.2
maxX <- minX + sqrt(areaKm2*1e6)
minY <- 7438877-1.6e5
maxY <- minY + sqrt(areaKm2*1e6)
meanY <- mean(c(minY, maxY))

# Add random noise to polygon
set.seed(5567913)
xAdd <- round(runif(1, -5e5, 1.5e6))
yAdd <- round(runif(1, 1e5, 5e5)) - xAdd/2
nPoints <- 20
betaPar <- 0.6
X = c(jitter(sort(rbeta(nPoints, betaPar, betaPar)*(maxX-minX)+minX)),
      jitter(sort(rbeta(nPoints, betaPar, betaPar)*(maxX-minX)+minX, decreasing = TRUE)))
Y = c(jitter(sort(rbeta(nPoints/2, betaPar, betaPar)*(maxY-meanY)+meanY)),
      jitter(sort(rbeta(nPoints, betaPar, betaPar)*(maxY-minY)+minY, decreasing = TRUE)),
      jitter(sort(rbeta(nPoints/2, betaPar, betaPar)*(meanY-minY)+minY)))

inputMapPolygon <- Polygon(cbind(X+xAdd, Y+yAdd)) %>%  list %>% 
  Polygons("s1") %>%  list %>%  SpatialPolygons(1L)
crs(inputMapPolygon) <- lcc05CRS
```

Module groups make loading multiple modules easier: only the name of the module group needs to be specified in the `simInit` call, which will then initialize the simulation with the child modules.

```{r module-group-init, eval=FALSE}
# setup simulation
outputDir <- file.path(tempdir(), "simOutputs")
times <- list(start=2005.0, end=2020.0)
parameters <- list(
  .globals = list(burnStats="nPixelsBurned", rnorm(1)),
  .progress = list(type="text", interval=1),
  cropReprojectLccAge = list(useCache=TRUE),
  forestSuccessionBeacons = list(returnInterval=1, startTime=times$start,
                                 .plotInitialTime=times$start, .plotInterval=1),
  forestAge = list(returnInterval=1, startTime=times$start+0.5,
                   .plotInitialTime=times$start, .plotInterval=1),
  fireSpreadLcc = list(nFires= 3, its=1e6, drought=1.2,
                       persistprob=0, returnInterval=1, startTime=times$start+1,
                       .plotInitialTime=times$start, .plotInterval=1),
  caribouMovementLcc = list(N=1e3, moveInterval=1, startTime=times$start+1,
                            torus=TRUE, glmInitialTime = NA_real_,
                            .plotInitialTime=times$start, .plotInterval=1)
)
modules <- list("LCC2005")
objects <- list(age="age", lcc05="lcc05", inputMapPolygon="inputMapPolygon")
paths <- list(
  cachePath = file.path(outputDir, "cache"),
  modulePath = "~/GitHub/McIntire-lab/modules",
  inputPath = inputDir,
  outputPath = outputDir
)

mySim <- simInit(times=times, params=parameters, modules=modules,
                 objects=objects, paths=paths)

modules(mySim) # note the child modules are initialized
```

Module groups can also be used to simplify the download of multiple modules:

```{r module-group-dl, eval=FALSE}
downloadModule("SpaDES_sampleModules") # not yet implemented
```
