---
title: "establishment"
author: ""
date: "15 November 2016"
output: pdf_document
---

# Overview

This is a very simple 'spread and establish' module. It has a contagious component, it evaluates conditions in a set of "starting" locations on a raster and a set of "potential arriving locations". This could emulate dispersal of many types of species, where the abundance of a set of locations determines how many "things" (e.g., seeds or insects) leave, and the "quality" of a set of locations which determines how many "things" land and establish.


## Teaching

This is currently used in teaching.

# Usage

```{r module_usage}

library(igraph)
library(SpaDES)
setPaths(modulePath="~/GitHub/McIntire-lab/modules")
library(ggplot2)

modules <- list("establishment")

mySim <- simInit(modules = modules)

# make sure the plotting device is clear
clearPlot()
spades(mySim)
```

# Experiment

What happens when we vary the number of source locations and the threshold for establishment?

```{r}

# factorial experiment
expList <- list(establishThresh=as.list(1:9/10), N = as.list(1:10*2))

# Run the experiment
set.seed(234)
dev()
clearPlot()
out <- experiment(mySim, params = list(establishment = expList),
                  .plotInitialTime=NA) # turn off plotting

# plot the experiment
expDF <- cbind(expand.grid(expList), 
               numEstablishments=
                 (sapply(out, function(sim) sum(sim$establish[]))))
expDF <- data.frame(sapply(expDF, unlist, recursive = FALSE))


g <- ggplot(expDF, aes(establishThresh, N)) +
  geom_raster(aes(fill=numEstablishments)) +
  scale_fill_gradient(high="darkgreen", low="lightgreen", trans = "log",
                      breaks=c(1,10,100,500, 1000)) 
  

g
  

```


## Plotting

The plot is simply a map of where establishment will occur.


# Data dependencies

## Input data

None required. This is a teaching module that creates its own inputs.

## Output data

Description of the module outputs.

# Links to other modules

Describe any anticipated linkages to other modules.

