---
title: "establishment"
author: ''
date: "7 December 2016"
output:
  html_document: default
  pdf_document: default
---

# Overview

This is a very simple 'spread and establish' module. It has a contagious component, it evaluates conditions in a set of "starting" locations on a raster and a set of "potential arriving locations". This could emulate dispersal of many types of species, where the abundance of a set of locations determines how many "things" (e.g., seeds or insects) leave, and the "quality" of a set of locations which determines how many "things" land and establish.



**This is currently used in teaching.**

# Usage

```{r module_usage, eval=TRUE, echo=TRUE, message=FALSE}

library(SpaDES)
modulePath <- file.path(dirname(tempdir()), "modules")
downloadModule("establishment", path = modulePath)
setPaths(modulePath=modulePath)

modules <- list("establishment")

mySim <- simInit(modules = modules)

# make sure the plotting device is clear
clearPlot()
spades(mySim)
```

## Plotting

The main plot is simply a map of where establishment will occur.


# Experiment

What happens when we vary the number of source locations and the threshold for establishment?

```{r experiment, eval=FALSE, cache=TRUE, echo=FALSE}

# factorial experiment
expList <- list(establishThresh=as.list(1:9/10), 
                Nsource = as.list(1:10*2))

# Run the experiment
set.seed(234)
if(interactive()) dev()
clearPlot()
out <- experiment(mySim, params = list(establishment = expList),
                  .plotInitialTime=NA) # turn off plotting

# plot the experiment -- how many establish by parameter combo
expDF <- cbind(expand.grid(expList), 
               numEstablishments=
                 (sapply(out, function(sim) sum(sim$establish[]))))
expDF <- data.frame(sapply(expDF, unlist, recursive = FALSE))


library(ggplot2)
g <- ggplot(expDF, aes(establishThresh, Nsource)) +
  geom_raster(aes(fill=numEstablishments)) +
  scale_fill_gradient(high="darkgreen", 
                      low="lightgreen", trans = "log",
                      breaks=c(1,10,100,500, 1000)) 
  

g
  

```



# Data dependencies

## Input data

None required. This is a teaching module that creates its own inputs.

## Output data

A binary raster layer indicating places that were both "close" to a source, and of high enough quality to pass the establishment threshold parameter.

